---
title: "AACT database sumary"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,comment = "")
```

## Introduction

The [Clinical Trials Transformatiion Initiative's](https://www.ctti-clinicaltrials.org/who-we-are/strategic-plan) mission is to "To develop and drive adoption of practices that will increase the quality and efficiency of clinical trials".

One of their projects is the [Aggregated Content of Clinical Trials](https://aact.ctti-clinicaltrials.org/), which is "AACT is a publicly available relational database that contains all information (protocol and result data elements) about every study registered in ClinicalTrials.gov."

The purpose if this notebook is to simply collect, display, and provide a a brief descriptive summary of the AACT tables. 

## List tables in database

Here I am connecting to the AACT database and just listing the tables to be investigated in this notebook.

```{r}

if(require("RPostgreSQL")){library(RPostgreSQL)}else{install.packages("RPostgreSQL");library(RPostgreSQL)}
if(require("DBI")){library(DBI)}else{install.packages("DBI");library(DBI)}
if(require("tidyverse")){library(tidyverse)}else{install.packages("tidyverse");library(tidyverse)}
if(require("skimr")){library(skimr)}else{devtools::install_github("ropenscilabs/skimr");library(skimr)}

drv <- dbDriver('PostgreSQL')

con <- dbConnect(drv, 
                 dbname="aact",
                 host="aact-db.ctti-clinicaltrials.org", 
                 port=5432,
                 user=readr::read_tsv(".my.aact.cnf")$u, 
                 password=readr::read_tsv(".my.aact.cnf")$pw
                 )

dbTables <- dbListTables(con)

dbTables

```

## Collect AACT tables

Here I'm looping through the tables and collecting their contents from the AACT database.

```{r}

tables <- list()
for(i in 1:length(dbTables)){
  try(tables[[dbTables[i]]] <- tbl(con,dbTables[i]) %>% collect())
}

```

## Display and skim populated AACT tables

```{r}
dne <- c()
for(name in dbTables){
  tab <- tables[[name]]
  if(!is.null(tab)){
    cat(paste0(name," table exists\n\n"))
    print(head(tab))
    print(skim(tab))
    cat("\n")
  }else{
    cat(paste0("table ",name," doesn't exist\n\n"))
    dne <- c(dne,name)
  }
}

if(length(dne)>0){
  cat("These tables\n",dne,"\n aren't populated in AACT")
  }else{
  cat("All tables are populated!")
  }
```


<!-- ## Brief overview of target tables -->

<!-- In this section, I'm giving brief points and showing a bit of the data from each table.  -->

<!-- ```{r} -->
<!-- for(name in names(target_tables)){ -->
<!-- tab <- tables[[name]] -->
<!-- if(nrow(tab)>0){ -->
<!--   cat(paste0(name," table exists\t")) -->
<!--   head(tab) -->
<!-- }else{ -->
<!--   cat(paste0("table ",name," doesn't exist\t")) -->
<!-- } -->
<!-- } -->

<!-- ``` -->

<!-- ### outcome analyses -->

<!-- ```{r} -->
<!-- tab <- tables[["outcome_analyses"]] -->
<!-- nrow(tab) -->
<!-- head(tab) -->
<!-- apply(tab,2,skimr::n_missing) -->
<!-- ``` -->

<!-- This table contains `r nrow(tab)` outcome analyses from `r nrow(tab)` clinical trials. I'm not sure what each column means...Luckily, this [document](ClinicalTrials.gov Results Data Element...terventional and Observational Studies) contains that information! The different columns list the types of statistical tests used and their parameters. Alot of these parameters are missing. -->

<!-- ### facility investigators -->

<!-- ```{r} -->
<!-- tab <- tables[["facility_investigators"]] -->
<!-- nrow(tab) -->
<!-- head(tab) -->
<!-- apply(tab,2,skimr::n_missing) -->
<!-- ``` -->

<!-- This table contains `r nrow(tab)` facility roles for `r nrow(tab)` clinical trials.  -->

<!-- ```{r} -->
<!-- tab %>% group_by(role) %>% count() %>% ggplot() + geom_bar(stat="identity",aes(role,n,fill=role)) + coord_flip() -->
<!-- ``` -->

<!-- Apparently, mostly principal investigators are listed and then sub-investigators. But no study chairs.  -->

<!-- ### outcome counts -->

<!-- ```{r} -->
<!-- tab <- tables[["outcome_counts"]] -->
<!-- nrow(tab) -->
<!-- head(tab) -->
<!-- apply(tab,2,skimr::n_missing) -->
<!-- ``` -->

<!-- This table contains the counts of outcomes, about `r nrow(tab)`, for `r length(unique(tab$nct_id))` clinical trials. This table is completely filled but only has information on a relatively few clinical trials. -->

<!-- ### drop withdrawals -->

<!-- ```{r} -->
<!-- tab <- tables[["drop_withdrawals"]] -->
<!-- nrow(tab) -->
<!-- head(tab) -->
<!-- apply(tab,2,skimr::n_missing) -->
<!-- ``` -->

<!-- This table contains study drop information for `r length(unique(tab$nct_id))` clinical trials. This table is completely filled but only has information on a relatively few clinical trials. -->

<!-- ###baseline measurements -->

<!-- ```{r} -->
<!-- tab <- tables[["baseline_measurements"]] -->
<!-- nrow(tab) -->
<!-- head(tab) -->
<!-- apply(tab,2,skimr::n_missing) -->
<!-- ``` -->

<!-- This table contains study measurement info for `r length(unique(tab$nct_id))` clinical trials. This table is missing some parameter values and only has information on a relatively few clinical trials. -->

<!-- ### outcome analysis groups -->

<!-- ```{r} -->
<!-- tab <- tables[["outcome_analysis_groups"]] -->
<!-- nrow(tab) -->
<!-- head(tab) -->
<!-- apply(tab,2,skimr::n_missing) -->
<!-- ``` -->

<!-- This table contains study outcome analysis information for `r length(unique(tab$nct_id))` clinical trials. This table is completely filled but only has information on a relatively few clinical trials. And I'm not sire what the ctgov_group_code means... -->

<!-- ### baseline counts -->

<!-- ```{r} -->
<!-- tab <- tables[["baseline_counts"]] -->
<!-- nrow(tab) -->
<!-- head(tab) -->
<!-- apply(tab,2,skimr::n_missing) -->
<!-- ``` -->

<!-- This table contains study baseline counts information for `r length(unique(tab$nct_id))` clinical trials. This table is completely filled but only has information on a relatively few clinical trials. I'm also not sure what these columnns mean... -->

<!-- ### browse interventions -->

<!-- ```{r} -->
<!-- tab <- tables[["browse_interventions"]] -->
<!-- nrow(tab) -->
<!-- head(tab) -->
<!-- apply(tab,2,skimr::n_missing) -->
<!-- ``` -->

<!-- This table contains study baseline counts information for `r length(unique(tab$nct_id))` clinical trials. This table is completely filled but only has information on a relatively few clinical trials. I'm also not sure what these columnns mean... -->

<!-- ### browse interventions -->

<!-- ```{r} -->
<!-- tab <- tables[["browse_interventions"]] -->
<!-- nrow(tab) -->
<!-- head(tab) -->
<!-- apply(tab,2,skimr::n_missing) -->
<!-- ``` -->

<!-- This table contains study baseline counts information for `r length(unique(tab$nct_id))` clinical trials. This table is completely filled but only has information on a relatively few clinical trials. I'm also not sure what these columnns mean... -->

## Notes and tables of interest

The primary key to link  the tables in this database is _nct\_id_

The _study\_references_ table might be interesting to use since it gives the PMID number for that clinical trial. The corresponding text can be analyzed to discover topics and trends amongst clinical trials, using methods like Latent Direchlet Allocation. Or, it might be interesting to analyze the missing citations to clinical trials-looks like there's over 20K publication recordings missing.

In _responsible\_parties_ it might be interesting to see who conducts these trials and what that distribution is like. Though, about 27500 clinical trials don't have this attribute. 

The _design\_outcomes_ table doesn't seem to be documented well...we may want to investigate this table a bit more.

The _ipd\_information\_types_ table shows the types of clinical trials. There's on about 6.5K names here.

The _overall\_officials_ table might be interesting to look at in combination with the _responsible\_parties_ table. There should be good agreement. 

The _designs\_table_ looks at the experimental design for the clinical trial, like if it was an observational study or the methodology used.

The _keywords_ table might be interesting to look at. I wonder how similar they are to MeSH terms?

The _participant\_flows_ table is kind of different-it seems to give a text description of the trial recruitment and before patients were assigned their study arm to be in. 

The _browse\_interventions_ table seems to be similar to the _keywords_ table but here they're labeled as MeSH terms.

The _facilities_ table would be good to look at to find out the location of these trials.

The _milestones_ table is more fully described [here](https://aact.ctti-clinicaltrials.org/points_to_consider). For the ctgov_group_code, those with P1 are 'All Milestones & Drop_Withdrawals associated with this study's experimental group link to this row.' and P2 are 'All Milestones & Drop_Withdrawals associated with this study's studies control group link to this row.'. This is one of the four results tables.

The _result\_agreements_ table is another results table but not about groups. 

The _calculated\_values_ table seems to be a summary table, but not sure of what...

The _facility\_contacts_ table seems just to have contact info for the trial's facilities.

The _baseline\_counts_ table is another results table that I think tells you of the number of participants in the control and experimental groups. Might be interesting to get the number of participants starting out in a trial.

The _mesh\_terms_ table gives more details on the mesh terms. 

The _design\_group\_interventions_ table just contains concept ids. This table links to the _intervention\_other\_names_ table for the brand name drug or device, it seems. 




