---
title: "AACT database sumary"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,comment = "")
```

## Introduction

The [Clinical Trials Transformatiion Initiative's](https://www.ctti-clinicaltrials.org/who-we-are/strategic-plan) mission is to "To develop and drive adoption of practices that will increase the quality and efficiency of clinical trials".

One of their projects is the [Aggregated Content of Clinical Trials](https://aact.ctti-clinicaltrials.org/), which is "AACT is a publicly available relational database that contains all information (protocol and result data elements) about every study registered in ClinicalTrials.gov."

The purpose if this notebook is to simply collect, display, and provide a a brief descriptive summary of the AACT tables. 

## List tables in database

Here I am connecting to the AACT database and just listing the tables to be investigated in this notebook.

```{r}

if(require("RPostgreSQL")){library(RPostgreSQL)}else{install.packages("RPostgreSQL");library(RPostgreSQL)}
if(require("DBI")){library(DBI)}else{install.packages("DBI");library(DBI)}
if(require("tidyverse")){library(tidyverse)}else{install.packages("tidyverse");library(tidyverse)}
if(require("skimr")){library(skimr)}else{devtools::install_github("ropenscilabs/skimr");library(skimr)}

drv <- dbDriver('PostgreSQL')

con <- dbConnect(drv, 
                 dbname="aact",
                 host="aact-db.ctti-clinicaltrials.org", 
                 port=5432,
                 user=readr::read_tsv(".my.aact.cnf")$u, 
                 password=readr::read_tsv(".my.aact.cnf")$pw
                 )

dbTables <- dbListTables(con)

dbTables

```

## Collect AACT tables

Here I'm looping through the tables and collecting their contents from the AACT database.

```{r}

tables <- list()
for(i in 1:length(dbTables)){
  try(tables[[dbTables[i]]] <- tbl(con,dbTables[i]) %>% collect())
}

```

## Display and skim populated AACT tables

```{r}
dne <- c()
for(name in dbTables){
  tab <- tables[[name]]
  if(!is.null(tab)){
    cat(paste0(name," table exists\n\n"))
    print(head(tab))
    print(skim(tab))
    cat("\n")
  }else{
    cat(paste0("table ",name," doesn't exist\n\n"))
    dne <- c(dne,name)
  }
}

if(length(dne)>0){
  cat("These tables\n",dne,"\n aren't populated in AACT")
  }else{
  cat("All tables are populated!")
  }
```


<!-- ## Brief overview of target tables -->

<!-- In this section, I'm giving brief points and showing a bit of the data from each table.  -->

<!-- ```{r} -->
<!-- for(name in names(target_tables)){ -->
<!-- tab <- tables[[name]] -->
<!-- if(nrow(tab)>0){ -->
<!--   cat(paste0(name," table exists\t")) -->
<!--   head(tab) -->
<!-- }else{ -->
<!--   cat(paste0("table ",name," doesn't exist\t")) -->
<!-- } -->
<!-- } -->

<!-- ``` -->

<!-- ### outcome analyses -->

<!-- ```{r} -->
<!-- tab <- tables[["outcome_analyses"]] -->
<!-- nrow(tab) -->
<!-- head(tab) -->
<!-- apply(tab,2,skimr::n_missing) -->
<!-- ``` -->

<!-- This table contains `r nrow(tab)` outcome analyses from `r nrow(tab)` clinical trials. I'm not sure what each column means...Luckily, this [document](ClinicalTrials.gov Results Data Element...terventional and Observational Studies) contains that information! The different columns list the types of statistical tests used and their parameters. Alot of these parameters are missing. -->

<!-- ### facility investigators -->

<!-- ```{r} -->
<!-- tab <- tables[["facility_investigators"]] -->
<!-- nrow(tab) -->
<!-- head(tab) -->
<!-- apply(tab,2,skimr::n_missing) -->
<!-- ``` -->

<!-- This table contains `r nrow(tab)` facility roles for `r nrow(tab)` clinical trials.  -->

<!-- ```{r} -->
<!-- tab %>% group_by(role) %>% count() %>% ggplot() + geom_bar(stat="identity",aes(role,n,fill=role)) + coord_flip() -->
<!-- ``` -->

<!-- Apparently, mostly principal investigators are listed and then sub-investigators. But no study chairs.  -->

<!-- ### outcome counts -->

<!-- ```{r} -->
<!-- tab <- tables[["outcome_counts"]] -->
<!-- nrow(tab) -->
<!-- head(tab) -->
<!-- apply(tab,2,skimr::n_missing) -->
<!-- ``` -->

<!-- This table contains the counts of outcomes, about `r nrow(tab)`, for `r length(unique(tab$nct_id))` clinical trials. This table is completely filled but only has information on a relatively few clinical trials. -->

<!-- ### drop withdrawals -->

<!-- ```{r} -->
<!-- tab <- tables[["drop_withdrawals"]] -->
<!-- nrow(tab) -->
<!-- head(tab) -->
<!-- apply(tab,2,skimr::n_missing) -->
<!-- ``` -->

<!-- This table contains study drop information for `r length(unique(tab$nct_id))` clinical trials. This table is completely filled but only has information on a relatively few clinical trials. -->

<!-- ###baseline measurements -->

<!-- ```{r} -->
<!-- tab <- tables[["baseline_measurements"]] -->
<!-- nrow(tab) -->
<!-- head(tab) -->
<!-- apply(tab,2,skimr::n_missing) -->
<!-- ``` -->

<!-- This table contains study measurement info for `r length(unique(tab$nct_id))` clinical trials. This table is missing some parameter values and only has information on a relatively few clinical trials. -->

<!-- ### outcome analysis groups -->

<!-- ```{r} -->
<!-- tab <- tables[["outcome_analysis_groups"]] -->
<!-- nrow(tab) -->
<!-- head(tab) -->
<!-- apply(tab,2,skimr::n_missing) -->
<!-- ``` -->

<!-- This table contains study outcome analysis information for `r length(unique(tab$nct_id))` clinical trials. This table is completely filled but only has information on a relatively few clinical trials. And I'm not sire what the ctgov_group_code means... -->

<!-- ### baseline counts -->

<!-- ```{r} -->
<!-- tab <- tables[["baseline_counts"]] -->
<!-- nrow(tab) -->
<!-- head(tab) -->
<!-- apply(tab,2,skimr::n_missing) -->
<!-- ``` -->

<!-- This table contains study baseline counts information for `r length(unique(tab$nct_id))` clinical trials. This table is completely filled but only has information on a relatively few clinical trials. I'm also not sure what these columnns mean... -->

<!-- ### browse interventions -->

<!-- ```{r} -->
<!-- tab <- tables[["browse_interventions"]] -->
<!-- nrow(tab) -->
<!-- head(tab) -->
<!-- apply(tab,2,skimr::n_missing) -->
<!-- ``` -->

<!-- This table contains study baseline counts information for `r length(unique(tab$nct_id))` clinical trials. This table is completely filled but only has information on a relatively few clinical trials. I'm also not sure what these columnns mean... -->

<!-- ### browse interventions -->

<!-- ```{r} -->
<!-- tab <- tables[["browse_interventions"]] -->
<!-- nrow(tab) -->
<!-- head(tab) -->
<!-- apply(tab,2,skimr::n_missing) -->
<!-- ``` -->

<!-- This table contains study baseline counts information for `r length(unique(tab$nct_id))` clinical trials. This table is completely filled but only has information on a relatively few clinical trials. I'm also not sure what these columnns mean... -->
