---
title: "Analysis"
author: "Jason Baik"
date: "10/29/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load packages
```{r}
library(tidyverse)
if(require("RPostgreSQL")){library(RPostgreSQL)}else{install.packages("RPostgreSQL");library(RPostgreSQL)}
if(require("DBI")){library(DBI)}else{install.packages("DBI");library(DBI)}
if(require("tidyverse")){library(tidyverse)}else{install.packages("tidyverse");library(tidyverse)}

drv <- dbDriver('PostgreSQL')
con <- dbConnect(drv, 
                 dbname="aact",
                 host="aact-db.ctti-clinicaltrials.org", 
                 port=5432,
                 user=readr::read_csv("id_pw.csv")$id, 
                 password=readr::read_csv("id_pw.csv")$pw
                 )

dbListTables(con) %>% View("List of Tables")
```

### What Nick wants to analyze

* Latent Direchlet Allocation with MeSH terms from trials
* Assess the life cycle of a clinical trial
* Aq  ssess the investigation of drugs across clinical trials for different conditions

### Description of List of Tables / Columns
www.ctti-clinicaltrials.org/files/aact201603_comprehensive_data_dictionary_1.xlsx

### Github Repos

1) https://github.com/statwonk/aact 

* txt format files of AACT data

2) https://github.com/kchis/AACT-Sample-Graphs

* example graphs summarizing characteristics of interventional trials in ClinicalTrials.gov, 2008-2017

3) https://github.com/datasciences1/aact
4) https://github.com/Shou-Yu-YAN/AACT_database


### Prove AACT pulled data from clinicaltrials.org
```{r, eval = FALSE}
# id_information table from AACT
aact_nct_id <- tbl(con, "id_information") %>% 
  collect() %>% 
  select(nct_id) %>% 
  pull() %>% 
  unique()


# Name (NCT IDs) of all the XML files
test <- list.files(path = "data/AllPublicXML/", pattern = "^NCT",
           recursive = TRUE)
orig_nct_id <- substr(test, 13, nchar(test) - 4)
orig_nct_id <- orig_nct_id %>% unique()

# Intersection
intersect(aact_nct_id, orig_nct_id) %>% length()
```


### EDA
```{r}
# Observe the first 22 tables
tbl(con, "documents") %>% 
  collect() %>% 
  head()
```

eligibilities -> `minimum_age` is a quantitative variable
brief_summaries -> text values of summaries
facilities -> text description of institutions where study was held
browse_conditions -> `nct_id` / mesh terms
schema_migrations -> Does not exist
links -> `url` of study
result_contacts -> contact info for researchers
design_outcomes -> text description of outcomes
mesh_headings -> heading / subcategory
result_groups -> text descriptions (example: Subjects were administered with insulin glargine (IGlar: 100 U/mL) subcutaneously (s.c.) once daily for a duration of 26 weeks)
sponsors -> list of all the sponsors
designs -> design of experiment
milestones -> 
studies -> date values of dates related to study
detailed_descriptions -> more detailed descriptions...
documents -> `document_type` seems important / interesting

```{r}
# purrr
list_tables <- dbListTables(con)
# Take out schema_migrations table, which does not exist
list_tables <- list_tables[-9] 

list_tables_head <- map(list_tables, function(x) x %>% 
      tbl(con, .) %>% 
      collect() %>% 
      head())
```


