---
title: "Clinical trial measurements"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = F,comment = "")
```

# Introduction

This notebook will explore measurements recorded for clinical trials. The utility of these measurements will be evaluated

```{r,warning=F,message=FALSE}

library(tidyverse)

source("helper_functions.R")

```

# Clinical trial measurements

These are one of the four results measurements for clinical trials [Ref](https://aact.ctti-clinicaltrials.org/schema). And each trial has different result groups. 

```{r}

con <- aact_connector()

```


The Result_Groups table represents an aggregate list of all groups associated with these result types. All result tables (Outcomes, Outcome_Counts, Baseline_Measures, Reported_Events, etc.) relate to Result_Groups via the foreign key result_group_id

```{r}

result_groups <- get_table(con,"result_groups") %>% collect()

```

How many result types are there for trials?

```{r}
tmp <- result_groups %>% 
  select(nct_id,result_type) %>% 
  distinct() %>% 
  group_by(nct_id) %>% 
  count()

tmp

table(tmp$n)
```

Every trial here has atleast 2 result_types, where majority have all 4 and only a few hundred has only three. 

How many trials are represented here? Only `r scales::comma(length(unique(result_groups$nct_id)))`

How many groups are in the result_type's?

```{r}
tmp <- result_groups %>% 
  group_by(nct_id,result_type) %>% 
  count()

tmp

table(tmp$n)
```

The `n` indicates there's n groups for each result_type. There could be a lot!

Ok cool! Let's try to get deeper.

## Baseline Measurements

```{r}

baseline_measurements <- get_table(con,"baseline_measurements") %>% collect()

```

For a random nct like *NCT00000125*, let's get the baseline measurements.

```{r}

nct="NCT00000125"

baseline_measurements %>% 
  filter(nct_id==nct)


```

With this we can see the number of participants.

```{r,fig.width=12}

baseline_measurements %>% 
  filter(nct_id==nct) %>% 
  select(ctgov_group_code,classification,title,units,param_value_num) %>% 
  ggplot() +
  geom_bar(aes(forcats::fct_reorder(classification,param_value_num),param_value_num,fill=ctgov_group_code),stat="identity") +
  facet_wrap(.~ctgov_group_code) +
  scale_y_continuous(trans = "log10",labels = scales::number) +
  coord_flip() +
  xlab("Race/Ethnicity") +
  ylab("Number of participants") +
  theme_bw() +
  guides(fill=guide_legend(title = "Baseline group"))

```

Interesting - so, at least for the baseline measurements, this tells me how many people were in each group.

This is probably what all the other measurements describe.

<!-- ## Outcome Measurements -->

<!-- ```{r} -->

<!-- outcome_measurements <- get_table(con,"outcome_measurements") %>% collect() -->

<!-- ``` -->

<!-- ```{r} -->

<!-- ``` -->
<!-- ## Reported events -->

<!-- ```{r} -->

<!-- reported_events <- get_table(con,"reported_events") %>% collect() -->

<!-- ``` -->

<!-- ```{r} -->

<!-- ``` -->
<!-- ## Participant flows -->

<!-- ```{r} -->

<!-- participant_flows <- get_table(con,"participant_flows") %>% collect() -->

<!-- ``` -->

<!-- ```{r} -->

<!-- ``` -->