---
title: "Types of drugs tested in clinical trials"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Here, I'm investigating the types of drugs tested in drug intervention clinical trials. 

Often, drugs are mapped to identifiers so that ambiguity can be alleviated when identifying drugs. The identifiers belong to certain standard vocabulary. A well known and used vocabulary is RxNorm - this vocabulary distinguishes drugs based on their formulation type. Another vocabulary is the Anatomical Therapeutic Class (ATC) - this v ocabulary categories drugs based on their mechanism of action. 

For this notebook, I am going to map the drug names to vocabularies.

First, looks like I need to know whether a drug trial has a drug intervention - that's in the _interventions_ table.

I want to use my cleaned _study\_phase\_age\_elibility.csv_ file too - I'll join the two. 

```{r}
library(tidyverse)
source("helper_functions.R")
```
```{r}

cleaned_study_age_phase_data <- read_csv("study_phase_age_eligibility.csv")

con <- aact_connector()

interventions <- get_table(con,"interventions") %>% filter(intervention_type=="Drug") %>% collect()

joined <- left_join(cleaned_study_age_phase_data,interventions)

```

Now I need the RxNorm id of the drug. I don't think it's in the database right now...well, I can join by name for now. Not sure how useful that'll be.

```{r}

joined

concept <- read_tsv("../../Research/Projects/pediatrics/vocabulary_download_v5_{d62de2e4-9825-4917-8ccb-d92395a5778e}_1535207123450/CONCEPT.csv") %>% filter(vocabulary_id=="ATC" | vocabulary_id=="RxNorm")

joined_to_drug_id <- left_join(joined,concept,by=c("name"="concept_name"))
```

Are these drugs mapped to RxNorm or ATC vocabularies?

```{r}

table(joined_to_drug_id$vocabulary_id)

```

Many to RxNorm and ATC. 

For ATC encoded drugs, what subclass do they map to?

```{r}

joined_to_drug_id %>% 
        filter(vocabulary_id=="ATC") %>% 
        count(concept_class_id) %>% 
  arrange(desc(n))

```

ATC 4th or 5th, so the most specific ATC categories.


For RxNorm encoded drugs, what subclass do they map to?

```{r}

joined_to_drug_id %>% 
        filter(vocabulary_id=="RxNorm") %>% 
        count(concept_class_id) %>% 
  arrange(desc(n))

```

It ranges from the very specific drug ingredients to the brand or clinical drug forms.

Great. The issue is I don't know how faithful the mapping is since I'm joining by name. Which can be ambiguous.