---
title: "Identifying interesting questions/hypotheses about clinical trials"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

I'm not really sure what exactly to investigate right now, so I'm going to try out a few things and see what's worth pursuing.

```{r,echo=F}

library(tidyverse)
library(data.table)

source("helper_functions.R")

theme_set(theme_bw())

```

## Clinical trial characteristics

I'm going to look at the studfy characteristics of clinical trials using the _studies_ table since it has a lot of attributes that I think would be useful. After looking at my [summary notebook](https://github.com/jasonbaik94/clinical_trials/blob/master/aact_database_table_summary.md), there's data on the study type, the phase of a clinical trial, start and end of duration of the trial, why it was stopped, and others. 

First I load the _studies_ table and get the relevant variables:

```{r}

con <- aact_connector()

studies <- get_table(con,"studies") %>% select(nct_id,study_first_submitted_date,phase,
                                               completion_date,study_type,
                                               baseline_population,overall_status,
                                               why_stopped,is_fda_regulated_drug) %>% 
  collect() %>% distinct()

dbDisconnect(con)

studies

```

## Phases of clinical trials

I'm interested in the types of phases given for clinical trials

```{r}
studies %>% 
  ggplot() +
  geom_bar(aes(forcats::fct_infreq(phase),color=phase),fill="gray") +
  coord_flip() +
  ylab("Number of clinical trials") +
  xlab("") +
  theme(
    legend.position = "none"
  )
```

Three things:

1) Over 15K trials don't have a phase attributed, as either NA or N/A

2) There's a preceding "Phase " before the actual number

3) There's some trials that are combinations of phases. What does that mean?

## Clinical trial phase clean up

1) Separate phase wording from numbers

2) make separate 'early' column for early phase 1 distinction from phase 1

3) make combo phase trials as phase 1 + '_' + phase 2. This will make the whole column a character which I'm ok with-don't want to treat these as continuous or integer but as distinct entities

4) Combine trials without a phase as one NA value not NA or N/A

5) save as a data file (make a data/ directory)

```{r}

phases <- studies$phase

table(phases) 

phase_splt <- strsplit(phases," ") 

nsplits <- sapply(phase_splt,length)

table(nsplits)

```

now I split phases by a space, and so either I get no splits for NA and N/A, 1 split for Phase [1234], or two splits for Early Phase 1 and Phase [12]/Phase [23].

For the no splits, I need to use one common identifier, NA

```{r}
phases[nsplits==1] <- NA
```

For 1 split, I need to use just the integer.

```{r}
table(phases[nsplits==2])

n2_phase_splt <- strsplit(phases[nsplits==2]," ")

phases[nsplits==2] <- sapply(n2_phase_splt,function(x){x[2]})
```

For 2 splits, I need to be more crafty. Let's first deal with early phase 1.

```{r}
earlies <- grepl("Early",phases)

table(phases[earlies]) #just checking

studies <- 
  studies %>% 
  mutate(early = earlies)

phases[earlies] <- "1"

```

Now combo trials.

```{r}

combos <- !earlies & nsplits==3

table(phases[combos]) #just checking

tmp <- strsplit(phases[combos]," ")

second <- sapply(tmp,function(x){x[3]})

tmp <- sapply(tmp,function(x){x[2]})

first <- sapply(strsplit(tmp,"/"),function(x){x[1]})

phases[combos] <- paste(first,second,sep="_")
```

Ok let's see the cleaned version

```{r}

table(phases)

studies <- studies %>% mutate(clean_phase = phases)

studies %>% 
  ggplot() +
  geom_bar(aes(forcats::fct_infreq(clean_phase),color=clean_phase),fill="gray") +
  coord_flip() +
  ylab("Number of clinical trials") +
  xlab("") +
  theme(
    legend.position = "none"
  )

```

That's better.

## Cleaning eligibility ages

Thev ages given to the eligibility are all coded differently, but I want them all in the same units

```{r}

con <- aact_connector()

eligibilities <- get_table(con,"eligibilities") %>% select(nct_id,nct_id,minimum_age,maximum_age) %>% 
  collect() %>% distinct()

dbDisconnect(con)

transform_trial_age_range <- function(df,...){
  
  df %>% 
    mutate(
      minimum_age_num = sapply(strsplit(minimum_age," "),function(x){x[1]}),
      minimum_age_unit = sapply(strsplit(minimum_age," "),function(x){x[2]}),
      maximum_age_num = sapply(strsplit(maximum_age," "),function(x){x[1]}),
      maximum_age_unit = sapply(strsplit(maximum_age," "),function(x){x[2]})
    )
  
}

eligibilities_sep <- transform_trial_age_range(eligibilities)

eligibilities_sep$minimum_age_num[eligibilities_sep$minimum_age_num=="N/A"] <- NA
eligibilities_sep$minimum_age_unit[eligibilities_sep$minimum_age_unit=="N/A"] <- NA
eligibilities_sep$maximum_age_num[eligibilities_sep$maximum_age_num=="N/A"] <- NA
eligibilities_sep$maximum_age_unit[eligibilities_sep$maximum_age_unit=="N/A"] <- NA

eligibilities_sep <- eligibilities_sep %>% select(-maximum_age,-minimum_age)

table(eligibilities_sep$minimum_age_unit)

eligibilities_sep <- eligibilities_sep %>% 
  rename(
    maximum_age = maximum_age_num,
    minimum_age = minimum_age_num
  )

make_master_eligibility_ages <- function(ct){
  
  round_n <- 0
  
  ct$minimum_age <- as.integer(ct$minimum_age)
  ct$maximum_age <- as.integer(ct$maximum_age)
  
  ct %>% 
  
  mutate(
    
    min_minute_years = ifelse(replace_na(stringr::str_detect(ct$minimum_age_unit,"Minute|Minutes"),F),round(minimum_age/60/24/7/52,round_n),minimum_age),
    min_hour_years = ifelse(replace_na(stringr::str_detect(ct$minimum_age_unit,"Hour|Hours"),F),round(minimum_age/24/7/52,round_n),minimum_age),
    min_day_years = ifelse(replace_na(stringr::str_detect(ct$minimum_age_unit,"Day|Days"),F),round(minimum_age/365,round_n),minimum_age),
    min_week_years = ifelse(replace_na(stringr::str_detect(ct$minimum_age_unit,"Week|Weeks"),F),round(minimum_age/52,round_n),minimum_age),
    min_month_years = ifelse(replace_na(stringr::str_detect(ct$minimum_age_unit,"Month|Months"),F),round(minimum_age/12,round_n),minimum_age),
    min_year_years = ifelse(replace_na(stringr::str_detect(ct$minimum_age_unit,"Year|Years"),F),minimum_age,minimum_age),
    
    min_minutes=as.integer(replace_na(stringr::str_detect(ct$minimum_age_unit,"Minute|Minutes"),F)),
    min_hours=as.integer(replace_na(stringr::str_detect(ct$minimum_age_unit,"Hour|Hours"),F)),
    min_days=as.integer(replace_na(stringr::str_detect(ct$minimum_age_unit,"Day|Days"),F)),
    min_weeks=as.integer(replace_na(stringr::str_detect(ct$minimum_age_unit,"Week|Weeks"),F)),
    min_months=as.integer(replace_na(stringr::str_detect(ct$minimum_age_unit,"Month|Months"),F)),
    min_years=as.integer(replace_na(stringr::str_detect(ct$minimum_age_unit,"Year|Years"),F)),
    
    max_minute_years = ifelse(replace_na(stringr::str_detect(ct$maximum_age_unit,"Minute|Minutes"),F),round(maximum_age/60/24/7/52,round_n),maximum_age),
    max_hour_years = ifelse(replace_na(stringr::str_detect(ct$maximum_age_unit,"Hour|Hours"),F),round(maximum_age/24/7/52,round_n),maximum_age),
    max_day_years = ifelse(replace_na(stringr::str_detect(ct$maximum_age_unit,"Day|Days"),F),round(maximum_age/365,round_n),maximum_age),
    max_week_years = ifelse(replace_na(stringr::str_detect(ct$maximum_age_unit,"Week|Weeks"),F),round(maximum_age/52,round_n),maximum_age),
    max_month_years = ifelse(replace_na(stringr::str_detect(ct$maximum_age_unit,"Month|Months"),F),round(maximum_age/12,round_n),maximum_age),
    max_year_years = ifelse(replace_na(stringr::str_detect(ct$maximum_age_unit,"Year|Years"),F),maximum_age,maximum_age),
    
    max_minutes=as.integer(replace_na(stringr::str_detect(ct$maximum_age_unit,"Minute|Minutes"),F)),
    max_hours=as.integer(replace_na(stringr::str_detect(ct$maximum_age_unit,"Hour|Hours"),F)),
    max_days=as.integer(replace_na(stringr::str_detect(ct$maximum_age_unit,"Day|Days"),F)),
    max_weeks=as.integer(replace_na(stringr::str_detect(ct$maximum_age_unit,"Week|Weeks"),F)),
    max_months=as.integer(replace_na(stringr::str_detect(ct$maximum_age_unit,"Month|Months"),F)),
    max_years=as.integer(replace_na(stringr::str_detect(ct$maximum_age_unit,"Year|Years"),F)),
    
    minimum_master_age = min_minute_years * min_minutes + min_hour_years * min_hours + min_day_years * min_days + min_week_years * min_weeks + min_month_years * min_months + min_year_years * min_years,
    maximum_master_age = max_minute_years * max_minutes + max_hour_years * max_hours + max_day_years * max_days + max_week_years * max_weeks + max_month_years * max_months + max_year_years * max_years
    
  )  %>% 
  data.table

}

eligibilities_sep_master_dt <- make_master_eligibility_ages(eligibilities_sep) %>% select(nct_id,minimum_master_age,maximum_master_age)

```

## Joining eligibility age and phases

I'm interested to see who predominantly enters certain trial phases.

```{r}

setkey(eligibilities_sep_master_dt,nct_id)

studies_dt <- data.table(studies)

setkey(studies_dt,nct_id)

study_eligibility_dt <- studies_dt[eligibilities_sep_master_dt][,.(nct_id,clean_phase,minimum_master_age,maximum_master_age)]

head(study_eligibility_dt)

```

output this for use later

```{r}

study_eligibility_dt %>% write_csv("study_phase_age_eligibility.csv")

```