---
title: "Analysis V2"
author: "Jason Baik"
date: "1/6/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Import
```{r, include=FALSE}
library(tidyverse)
library(tidytext)
library(topicmodels)
library(RPostgreSQL)
library(DBI)
library(broom)
library(readxl)
library(tm)
theme_set(theme_light())

# Connect to PostgreSQL database
drv <- dbDriver('PostgreSQL')
con <- dbConnect(drv, 
                 dbname="aact",
                 host="aact-db.ctti-clinicaltrials.org", 
                 port=5432,
                 user=readr::read_csv("id_pw.csv")$id, 
                 password=readr::read_csv("id_pw.csv")$pw
                 )

# Import datatable of MeSH terms called browse_conditions
browse_conditions <- tbl(con, "browse_conditions") %>% 
  collect()
```


### Try joining on lower case MeSH terms

```{r}
tagged_mesh_terms_lowercase <- read_excel("2010_tagged_mesh_and_free_text_terms.xlsx",
                                          sheet = 3) %>%
  janitor::clean_names() %>% 
  select(-c(term_type, identifier)) %>% 
  distinct()
```


## Data Prep before LDA

1) Take out comma and word after comma. (`str_replace`) for both `browse_conditions` and `tagged_mesh_terms_lowercase`
2) `inner_join` both datasets
3) Run `distinct` on merged dataset

```{r}
tagged_mesh_terms_lowercase_replaced <- tagged_mesh_terms_lowercase %>% 
  mutate(term = str_replace_all(term,
                                "(.*),.*",
                                "\\1"))

mesh <- browse_conditions %>% 
  mutate(downcase_mesh_term = str_replace_all(downcase_mesh_term,
                                              "(.*),.*", "\\1")) %>%
  inner_join(tagged_mesh_terms_lowercase_replaced, by = c("downcase_mesh_term" = "term")) %>%
  select(nct_id, downcase_mesh_term) %>% 
  distinct() 
```

## LDA

```{r}
mesh_count <- mesh %>% 
  count(nct_id, downcase_mesh_term, sort = TRUE) %>% 
  ungroup()

# Create DocumentTermMatrix in order to put dataset into the LDA algorithm from the topicmodels package
mesh_dtm <- mesh_count %>%
  cast_dtm(nct_id, downcase_mesh_term, n)

# This take a few minutes
mesh_lda <- topicmodels::LDA(mesh_dtm, k = 10, control = list(seed = 1234))

 # Use the broom package to mutate beta, the probability of that term being generated from that topic.
mesh_tidy <- tidy(mesh_lda)

# Use top_n() to find top 5 terms within each topic
mesh_top5 <- mesh_tidy %>% 
    group_by(topic) %>% 
    top_n(5, beta) %>% 
    ungroup() %>% 
    arrange(topic, -beta)

# Visualize mesh_top5
mesh_top5 %>%
  mutate(term = reorder(term, beta)) %>% 
  ggplot(aes(term, beta, fill = term)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free", nrow = 3) +
  coord_flip() +
  scale_y_continuous(labels = scales::percent_format()) +
  theme(axis.text.x = element_text(angle = 50, hjust = 1)) +
  labs(x = "MeSH",
       y = "Beta (Probability of MeSH being generated from topic)",
       caption = "@jsonbaik") +
  ggtitle("Distribution of MeSH")
```


Now comes the fun part: comparison between LDA's categorization of topics and Duke researchers' categorization (the ground truth)

Create a function and `map` over it.
```{r}
mesh_top5 %>% 
  filter(topic == 1) 

tagged_mesh_terms_lowercase_replaced %>% 
  count(clinical_domain, term, sort = TRUE) %>% 
  group_by(clinical_domain) %>% 
  mutate(ground_truth_beta_percentage = 100 * (n / sum(n))) %>% 
  select(-n) %>% 
  # Group by clinical_domain and then arrange by descending order of ground_truth_beta_percentage
  arrange(clinical_domain, -ground_truth_beta_percentage)
```









Problem: Some of these MeSH terms have multiple clinical domains tagged

```{r}
# For each clinical_domain, arrange terms by decreasing order of percentage_term
distribution_term <- tagged_mesh_terms %>%
  count(clinical_domain, term) %>% 
  group_by(clinical_domain) %>% 
  mutate(percentage_term = (n / sum(n)) * 100) %>% 
  arrange(clinical_domain, -percentage_term)
```

```{r}
# Top 5 terms for each clinical domain in terms of percentage_term 
top5_terms <- tagged_mesh_terms %>%
  count(clinical_domain, term) %>% 
  group_by(clinical_domain) %>% 
  mutate(percentage_term = (n / sum(n)) * 100) %>% 
  top_n(5, percentage_term) %>% 
  # Arrange percentage terms in decreasing order
  arrange(clinical_domain, -percentage_term)
```

```{r}
top5_terms %>% 
  filter(!(clinical_domain %in% c("oncology_general", "oncology_specific"))) %>% 
  ggplot(aes(x = term, y = percentage_term, fill = term)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  facet_wrap(~clinical_domain, nrow = 4, scales = "free")
```