---
title: 'Tagged Terms'
author: "Howard Baek"
date: "12/30/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and connect to PostgreSQL database 

```{r, warning=FALSE, message = FALSE}
library(tidyverse)
library(tidytext)
library(topicmodels)
library(RPostgreSQL)
library(DBI)
library(broom)
library(readxl)
theme_set(theme_light())
```


```{r}
drv <- dbDriver('PostgreSQL')
con <- dbConnect(drv, 
                 dbname="aact",
                 host="aact-db.ctti-clinicaltrials.org", 
                 port=5432,
                 user=readr::read_csv("id_pw.csv")$id, 
                 password=readr::read_csv("id_pw.csv")$pw
                 )

# browse_conditions
browse_conditions <- tbl(con, "browse_conditions") %>% 
  collect()
```



## Using Sheri's Tags
```{r}
# Import data
tagged_mesh_terms <- read_excel("2010_tagged_mesh_and_free_text_terms.xlsx",
                                sheet = 3) %>%
  janitor::clean_names() %>% 
  select(-c(term_type, identifier)) %>% 
  mutate(term = str_to_title(term))

# Find unique MeSH terms
tagged_mesh_terms %>% 
  select(term) %>% 
  unique() %>% 
  nrow() # 4070 rows
  
browse_conditions %>% 
  select(mesh_term) %>% 
  unique() %>% 
  nrow() # 3898 rows

tagged_mesh_terms %>% 
  count(term, clinical_domain, sort = TRUE)

tagged_mesh_terms %>% 
  filter(term == "Neurofibromatosis 2")
```

Join `tagged_mesh_terms` and `browse_conditions` and get number of unique `clinical_domain`
```{r}
browse_conditions %>%
  select(nct_id, mesh_term) %>% 
  anti_join(tagged_mesh_terms, by = c("mesh_term" = "term")) %>%
  select(mesh_term) %>% 
  unique()

# 2201 mesh_terms didn't join
browse_conditions %>%
  select(mesh_term) %>% 
  unique()
# 2201 out of 3898 mesh terms didn't join (56.4%)
```

Let's move on with mesh_terms that joined (44% of mesh_terms in browse_conditions)
Where # of Topic (Clinical Domains) = 10. 
```{r}
browse_conditions_filtered <- browse_conditions %>%
  select(nct_id, mesh_term) %>% 
  inner_join(tagged_mesh_terms, by = c("mesh_term" = "term")) %>% 
  select(-clinical_domain) 
```

Perform LDA with K = 10
```{r}
browse_conditions_filtered_counts <- browse_conditions_filtered %>% 
  count(nct_id, mesh_term, sort = TRUE) %>% 
  ungroup()

browse_conditions_dtm <- browse_conditions_filtered_counts %>% 
  cast_dtm(nct_id, mesh_term, n)

topic_lda <- topicmodels::LDA(browse_conditions_dtm, k = 10, control = list(seed = 1234))

topic_lda_tidy <- tidy(topic_lda)

top_mesh_terms <- topic_lda_tidy %>% 
    group_by(topic) %>% 
    top_n(5, beta) %>% 
    ungroup() %>% 
    arrange(topic, -beta)
```

Plot first 5 topics
```{r}
top_mesh_terms %>%
  filter(topic %in% c(1:5)) %>% 
  mutate(term = reorder(term, beta)) %>% 
  ggplot(aes(term, beta, fill = term)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free", nrow = 3) +
  coord_flip() +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(x = "MeSH",
       y = "Beta (Probability of MeSH being generated from topic)") +
  ggtitle("Distribution of MeSH for 10 Topics from LDA")
```

Compare LDA output to "Ground Truth"
```{r}
# MeSH terms in Topic 1
top_mesh_terms_topic1 <- top_mesh_terms %>% 
  filter(topic == 1) %>% 
  select(term) %>% 
  pull()


# Check Topic 1 (See if LDA output matches with the "Ground Truth")
tagged_mesh_terms %>%
  filter(term %in% top_mesh_terms_topic1)
```

* Topic 1: Oncology

```{r}
# MeSH terms in Topic 2
top_mesh_terms_topic2 <- top_mesh_terms %>% 
  filter(topic == 2) %>% 
  select(term) %>% 
  pull()

# Check Topic 2
tagged_mesh_terms %>%
  filter(term %in% top_mesh_terms_topic2) %>% 
  count(clinical_domain)
```

* Topic 2: Cardiology, Oncology, Otolaryngology

```{r}
# MeSH terms in Topic 3
top_mesh_terms_topic3 <- top_mesh_terms %>% 
  filter(topic == 3) %>% 
  select(term) %>% 
  pull()

# Check Topic 2
tagged_mesh_terms %>%
  filter(term %in% top_mesh_terms_topic3) %>% 
  count(clinical_domain)
```


* Topic 3: Oncology, General Psych

```{r}
# MeSH terms in Topic 4
top_mesh_terms_topic4 <- top_mesh_terms %>% 
  filter(topic == 4) %>% 
  select(term) %>% 
  pull()

# Check Topic 4
tagged_mesh_terms %>%
  filter(term %in% top_mesh_terms_topic4) %>% 
  count(clinical_domain, sort = TRUE)
```

* Topic 4: Otolaryngology, Oncology, Immuno Rheumatology

Okay...Enough with the copy and pasting. Let's create a function and use `map`
```{r}
check_ground_truth <- function(topic_number) {
  
  topic_term <- top_mesh_terms %>% 
    filter(topic == topic_number) %>% 
    select(term) %>% 
    pull()
  
tagged_mesh_terms %>%
  filter(term %in% topic_term) %>% 
  count(clinical_domain, sort = TRUE)
}
```

```{r}
ground_truth <- map(c(1:10), check_ground_truth)

ground_truth[[7]]
```

