---
title: "Analysis V2"
author: "Howard Baek"
date: "1/6/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Import
```{r, include=FALSE}
library(tidyverse)
library(tidytext)
library(topicmodels)
library(RPostgreSQL)
library(DBI)
library(broom)
library(readxl)
library(tm)
theme_set(theme_light())

# Connect to PostgreSQL database
drv <- dbDriver('PostgreSQL')
con <- dbConnect(drv, 
                 dbname="aact",
                 host="aact-db.ctti-clinicaltrials.org", 
                 port=5432,
                 user=readr::read_csv("id_pw.csv")$id, 
                 password=readr::read_csv("id_pw.csv")$pw
                 )

# Import datatable of MeSH terms called browse_conditions
browse_conditions <- tbl(con, "browse_conditions") %>% 
  collect()
```


### Try joining on lower case MeSH terms

```{r}
tagged_mesh_terms_lowercase <- read_excel("2010_tagged_mesh_and_free_text_terms.xlsx",
                                          sheet = 3) %>%
  janitor::clean_names() %>% 
  select(-c(term_type, identifier)) %>% 
  distinct()
```


## Data Prep before LDA

1) Take out comma and word after comma. (`str_replace`) for both `browse_conditions` and `tagged_mesh_terms_lowercase`
2) `inner_join` both datasets
3) Run `distinct` on merged dataset. Name it `mesh`

```{r}
tagged_mesh_terms_lowercase_replaced <- tagged_mesh_terms_lowercase %>% 
  mutate(term = str_replace_all(term,
                                "(.*),.*",
                                "\\1"))

mesh <- browse_conditions %>% 
  mutate(downcase_mesh_term = str_replace_all(downcase_mesh_term,
                                              "(.*),.*", "\\1")) %>%
  inner_join(tagged_mesh_terms_lowercase_replaced, by = c("downcase_mesh_term" = "term")) %>%
  select(nct_id, downcase_mesh_term) %>% 
  distinct() 
```

## LDA

```{r}
mesh_count <- mesh %>% 
  count(nct_id, downcase_mesh_term, sort = TRUE) %>% 
  ungroup()


# Create DocumentTermMatrix in order to put dataset into the LDA algorithm from the topicmodels package
mesh_dtm <- mesh_count %>%
  cast_dtm(nct_id, downcase_mesh_term, n)

# This take a few minutes
mesh_lda <- topicmodels::LDA(mesh_dtm, k = 10, control = list(seed = 1234))

# Use the broom package to mutate beta, the probability of that term being generated from that topic.
mesh_tidy <- tidy(mesh_lda)

# Use top_n() to find top 5 terms within each topic
mesh_top5 <- mesh_tidy %>% 
    group_by(topic) %>% 
    top_n(5, beta) %>% 
    ungroup() %>% 
    arrange(topic, -beta)

# Visualize mesh_top5
mesh_top5 %>%
  mutate(term = reorder(term, beta)) %>% 
  ggplot(aes(term, beta, fill = term)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free") +
  coord_flip() +
  scale_y_continuous(labels = scales::percent_format()) +
  theme(axis.text.x = element_text(angle = 50, hjust = 1)) +
  labs(x = "MeSH",
       y = "Beta (Probability of MeSH being generated from topic)",
       caption = "@jsonbaik") +
  ggtitle("Distribution of MeSH")
```

```{r}
mesh_tidy %>% 
  filter(term != "disease") %>% 
  spread(term, beta)
```


Compare to Ground Truth
```{r}
browse_conditions %>% 
  mutate(downcase_mesh_term = str_replace_all(downcase_mesh_term,
                                              "(.*),.*", "\\1")) %>%
  inner_join(tagged_mesh_terms_lowercase_replaced, by = c("downcase_mesh_term" = "term")) %>% 
  distinct()
```

# TODO: Compare ground truth to LDA algorithm. 

- I need to find a way to calculate Betas in of each mesh term in each clinical domain

 
 